---
title: "Thesis2022_PlantFamilies"
author: "Amanda Hastings"
date: "2022-11-08"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Read in data 
ThesisDataRaw <- read.csv("HastingsA_MasterThesisData8.8.22.csv")
TerraData <- read.csv("HastingsA_TerraclimateMaster8.16.csv")
ThesisSeedlingsData <- read.csv("HastingsA_MasterSeedlingsData9.12.22.csv")

#Load libraries
library(tidyverse)
library(dplyr)
library(RColorBrewer)
library(emmeans)
library(multcomp)
library(multcompView)
library(ggthemes)
library(vegan)
library(ggvegan)
library(ggpubr)
library(viridis)
library(scales)
```

## Formatting raw thesis site data and seedlings data 

```{r}
#Filter to only wanted sites in ThesisData 
ThesisData <- ThesisDataRaw %>% 
  filter(!PlotID %in% c('LS11','HS02','HS12','HS13', 'LS01_MV','LS02_MV','HS01_MV'))

ThesisSeedlings <- ThesisSeedlingsData %>% 
  filter(!PlotID %in% c('LS11','HS02','HS12','HS13'))%>% 
  dplyr::select(!c(Fire_name, regen_radius, regenplot_size_m2, regenplot_size_ha))%>% 
  select_if(~!any(is.na(.)))

```

## Combining site data with seedlings data

```{r}
ThesisAll1 <- full_join(ThesisData, ThesisSeedlings, by= "PlotID") %>% 
  arrange(PlotID)%>% 
  mutate(forest_type = factor(forest_type))
```

## Add in plant family data frame 

```{r}
#Load plant family data 
families_data <- read.csv("HastingsA_FamilyCounts_11.7.22.csv")

families_data <- families_data %>% 
  mutate("percent_cover" = Count/127)

n_distinct(families_data$PlotID)
```


## Append burn severity and forest type to PlotID

```{r}
PlotCat <- ThesisAll1 %>% 
  dplyr::select(PlotID, burn_sev_category, forest_type)%>% 
  filter(!PlotID %in% "LS30")

#Append categories to PlotID in plant family dataframe 
families_data <- left_join(families_data, PlotCat, by = "PlotID")

#Summarize plant family cover by burn severity and forest type 
families_sumstats <- families_data %>%
  group_by(forest_type, burn_sev_category, Family) %>%
  mutate(forest_type = factor(
    forest_type,
    levels = c("Lower Montane", "Upper Montane","Subalpine")))%>%
  summarize(mean_percent_cover = mean(percent_cover)) %>%
  mutate(mean_percent_cover = replace(mean_percent_cover, mean_percent_cover == 0.0000, NA))%>% 
  na.omit()%>% 
  filter(mean_percent_cover > 0.01)

families_means <- families_data %>% 
  group_by(Family)%>% 
  summarize (mean_family_allsites = mean(percent_cover))
```

## Plant families by burn severity  

```{r}

#Plant family plot by burn severity 
PlantFamily_Burn <- ggplot(
  families_sumstats,
  aes(x = Family,
      y = mean_percent_cover, 
      fill = burn_sev_category)
) +
  geom_boxplot() + 
  # ylim(0, 0.6) +
  labs(x = "Plant Family",
       y = "Percent Cover",
       fill = "Burn Severity") +
  theme(legend.position = "bottom",
        legend.key.width = unit(1, "cm")) +
  theme_few()+
  scale_fill_manual(
    name = "Burn Severity",
    labels = c("High", "Low"), values=c("#bd0026","#fd8d3c"))+
  scale_y_continuous(labels = percent)+
  theme(axis.text.x = element_text(
    color = "black",
    size = 7,
    angle = 90
  ),
  axis.title.x = element_text(size = 11),
  axis.title.y = element_text(size = 11),
  axis.text.y = element_text(size = 8),
  legend.title = element_text(size = 9), 
  legend.text = element_text(size =9))
  # scale_x_discrete(labels = c("Forbs", "Graminoids", "Shrubs", "Seedlings"))+
  

PlantFamily_Burn

```

## Plant families by forest type 


```{r}

#Plant family plot by forest type  
PlantFamily_Forest <- ggplot(
  families_sumstats,
  aes(x = Family,
      y = mean_percent_cover, 
      fill = forest_type)
) +
  geom_boxplot() + 
  labs(x = "Plant Family",
       y = "Percent Cover",
       fill = "Forest Type") +
  theme(legend.position = "bottom",
        legend.key.width = unit(1, "cm")) +
  theme_few()+
  scale_fill_manual(
    name = "Forest Type",
    labels = c("Lower Montane", "Upper Montane", "Subalpine"), values=c("#9ebc00","#ffc107","#981679"))+
  theme(axis.text.x = element_text(
    color = "black",
    size = 7,
    angle = 90
  ),
  axis.title.x = element_text(size = 11),
  axis.title.y = element_text(size = 11),
  axis.text.y = element_text(size = 8),
  legend.title = element_text(size = 9), 
  legend.text = element_text(size =9)) + 
  # scale_x_discrete(labels = c("Forbs", "Graminoids", "Shrubs", "Seedlings"))+
  scale_y_continuous(labels = percent)

PlantFamily_Forest

```


## Plant families by forest type with facet wrap

```{r}
#Plant family plot by forest type facet 
PlantFamily_ForestFacet <- ggplot(
  families_sumstats,
  aes(x = Family,
      y = mean_percent_cover, 
      fill=forest_type)
) +
  geom_boxplot(alpha =0.95) +
  facet_wrap(~ forest_type, scales="free_x") +
  # ylim(0, 0.6) +
  # theme_classic() +
  labs(x = "Plant Family",
       y = "Percent Cover") +
  theme(legend.position = "bottom",
        legend.key.width = unit(1, "cm")) +
  theme_few()+
  scale_fill_manual(
    name = "Forest Type",
    labels = c("Lower Montane", "Upper Montane", "Subalpine"), values=c("#9ebc00","#ffc107","#981679"))+
  theme(axis.text.x = element_text(
    color = "black",
    size = 7,
    angle = 90,
    vjust = 0.25
  ),
  axis.title.x = element_text(size = 11),
  axis.title.y = element_text(size = 11),
  axis.text.y = element_text(size = 8),
  legend.title = element_text(size = 9), 
  legend.text = element_text(size =9)) + 
  # scale_x_discrete(labels = c("Forbs", "Graminoids", "Shrubs", "Seedlings"))+
  scale_y_continuous(labels = percent)

PlantFamily_ForestFacet

```

## Separating by forest type 

```{r}

families_lowermontane <- families_sumstats %>% 
  filter(forest_type == "Lower Montane")


families_uppermontane <- families_sumstats %>% 
  filter(forest_type == "Upper Montane")


families_subalpine <- families_sumstats %>% 
  filter(forest_type == "Subalpine")
```


```{r}
#Plant family plot by forest type facet 
Subalpine_families <- ggplot(
  families_subalpine,
  aes(x = Family,
      y = mean_percent_cover)
) +
  geom_boxplot() +
  labs(x = "Plant Family",
       y = "Percent Cover") +
  theme(legend.position = "bottom",
        legend.key.width = unit(1, "cm")) +
  theme_few()+
  # scale_fill_manual(
  #   name = "Burn Severity",
  #   labels = c("High", "Low"), values=c("#bd0026","#fd8d3c"))+
  theme(axis.text.x = element_text(
    color = "black",
    size = 7,
    angle = 90,
    vjust = .8,
    hjust = .8
  ),
  axis.title.x = element_text(size = 11),
  axis.title.y = element_text(size = 11),
  axis.text.y = element_text(size = 8),
  legend.title = element_text(size = 9), 
  legend.text = element_text(size =9)) + 
  # scale_x_discrete(labels = c("Forbs", "Graminoids", "Shrubs", "Seedlings"))+
  scale_y_continuous(labels = percent)

Subalpine_families
```



```{r}
#Plant family plot by forest type facet 
Uppermontane_families <- ggplot(
  families_uppermontane,
  aes(x = Family,
      y = mean_percent_cover)
) +
  geom_boxplot() +
  labs(x = "Plant Family",
       y = "Percent Cover") +
  theme(legend.position = "bottom",
        legend.key.width = unit(1, "cm")) +
  theme_few()+
  # scale_fill_manual(
  #   name = "Burn Severity",
  #   labels = c("High", "Low"), values=c("#bd0026","#fd8d3c"))+
  theme(axis.text.x = element_text(
    color = "black",
    size = 7,
    angle = 90,
    vjust = .8,
    hjust = .8
  ),
  axis.title.x = element_text(size = 11),
  axis.title.y = element_text(size = 11),
  axis.text.y = element_text(size = 8),
  legend.title = element_text(size = 9), 
  legend.text = element_text(size =9)) + 
  # scale_x_discrete(labels = c("Forbs", "Graminoids", "Shrubs", "Seedlings"))+
  scale_y_continuous(labels = percent)

Uppermontane_families
```



```{r}
#Plant family plot by forest type facet 
Lowermontane_families <- ggplot(
  families_lowermontane,
  aes(x = Family,
      y = mean_percent_cover)
) +
  geom_boxplot() +
  labs(x = "Plant Family",
       y = "Percent Cover") +
  theme(legend.position = "bottom",
        legend.key.width = unit(1, "cm")) +
  theme_few()+
  # scale_fill_manual(
  #   name = "Burn Severity",
  #   labels = c("High", "Low"), values=c("#bd0026","#fd8d3c"))+
  theme(axis.text.x = element_text(
    color = "black",
    size = 7,
    angle = 90,
    vjust = .8,
    hjust = .8
  ),
  axis.title.x = element_text(size = 11),
  axis.title.y = element_text(size = 11),
  axis.text.y = element_text(size = 8),
  legend.title = element_text(size = 9), 
  legend.text = element_text(size =9)) + 
  # scale_x_discrete(labels = c("Forbs", "Graminoids", "Shrubs", "Seedlings"))+
  scale_y_continuous(labels = percent)

Lowermontane_families
```

## Plant families analyses 

```{r}
#Summarize plant family cover by burn severity and forest type 
families_sumstats2 <- families_data %>%
  group_by(forest_type, burn_sev_category, Family) %>%
  mutate(forest_type = factor(
    forest_type,
    levels = c("Lower Montane", "Upper Montane","Subalpine")))%>%
  summarize(mean_percent_cover = mean(percent_cover)) %>%
  mutate(mean_percent_cover = replace(mean_percent_cover, mean_percent_cover == 0.0000, NA))%>% 
  na.omit() %>% 
  filter(mean_percent_cover > 0.01)

hist(families_sumstats2$mean_percent_cover)
hist(sqrt(families_sumstats2$mean_percent_cover))
```


```{r}
#
LMFamilies <- lm(sqrt(mean_percent_cover) ~ Family, data = families_sumstats2)
anova(LMFamilies)
par(mfrow = c(1, 2))
plot(LMFamilies, which = 1:2)
```


```{r}
emoutFamilies <- emmeans(LMFamilies, ~ Family)
emoutFamilies
pairs(emoutFamilies)
cld(emoutFamilies)
```

## NMDS data frame plant families 

```{r}
#Data frame with family counts, pivot wider
families_NMDS <- families_data %>% 
  dplyr::select(PlotID, forest_type, burn_sev_category, Family, Count) %>% 
  filter(!Count ==0)%>% 
  na.omit()%>% 
  pivot_wider(names_from = Family, values_from = Count)

families_NMDS[is.na(families_NMDS)] <- 0

#standardize by total method 
families_NMDS_matrix <- families_NMDS %>% 
  dplyr:: select(-forest_type, -burn_sev_category)%>%
  remove_rownames %>% 
  column_to_rownames(var="PlotID")%>% 
  decostand(., method='total')

families.NMDS.nopred <- families_NMDS %>% 
  dplyr:: select(-forest_type, -burn_sev_category)%>%
  remove_rownames %>% 
  column_to_rownames(var="PlotID")

#Create matrix with predictor categories 
families_NMDS_pred <- families_NMDS %>% 
  dplyr:: select(PlotID, forest_type, burn_sev_category)%>%
  remove_rownames %>% 
  column_to_rownames(var="PlotID")
```

## Running NMDS

```{r}
library(ggvegan)
NMDS_families <- metaMDS(families_NMDS_matrix, distance = "bray", autotransform = F, k=2)
ordiplot(NMDS_families, type="t")


autoplot(NMDS_families)
  # scale_color_manual(
  #   name = "Burn Severity",
  #   labels = c("High", "Low"),
  #   values = c("#bd0026", "#fd8d3c")
  # )

```



## NMDS families plot 

```{r}
#Fortified ordination - plant families
fort_Families <- fortify(NMDS_families)

#Grab burn severity and forest type from plot ID
fort.families.pred <- families_NMDS %>% 
   dplyr::select(PlotID,forest_type, burn_sev_category)


#Fortified data frame plus burn severity and forest type to plot
fort.plot_Families <- merge(fort_Families, fort.families.pred, by.x = "Label", by.y = "PlotID", all = T)
  
```


## Adding burn severity and forest type categories 

```{r}
#Lower Montane env fit 
EnvFit_Families <- envfit(NMDS_families, families_NMDS_pred)
# plot(EnvFit_LowerMontane)
EnvFit_Families
```


```{r}
adonis2(families.NMDS.nopred ~burn_sev_category*forest_type, data = families_NMDS_pred, distance ="bray")
```


```{r}
# mrpp(families.NMDS.nopred)
```


```{r}
# Burn severity as factor
Families.env.scores <- as.data.frame(scores(EnvFit_Families, display = "factors"))


Families.env.scores <- cbind(Families.env.scores, Categories = rownames(Families.env.scores))
#    mutate("Categories2" = case_when(Categories == "forest_typeLowerMontane"  ~ "Lower Montane", 
# #                                 EnvPred == "heat_load" ~ "Heat Load", 
# #                                 EnvPred == "canopy_cover" ~ "Canopy Cover", 
# #                                 EnvPred == "conifer_seed_source" ~ "Distance to Seed Source", 
# #                                 EnvPred == "live_ba" ~ "Live Basal Area", 
# #                                 EnvPred == "organic_cover" ~ "Litter/Woody Debris", 
# #                                 EnvPred == "burn_sev_categoryHigh" ~ "High Severity Fire",
# #                                 EnvPred == "CWD" ~ "Coarse Woody Debris"))%>%

```


## Families plot with ggplot 

```{r}
library(ggrepel)
#Families NMDS plot
FamiliesBurnNMDSplot <- ggplot() +
  geom_point(
    data = subset(fort.plot_Families, Score == "sites"),
    size = 2,
    alpha = 0.75,
    shape = "circle",
    stroke = 1,
    mapping = aes(x = NMDS1, y = NMDS2, color = burn_sev_category)
  ) +
  scale_color_manual(
    name = "",
    labels = c("High Severity", "Low Severity"),
    values = c("#bd0026", "#fd8d3c")
  ) +
  # geom_segment(
  #   data = LowerMontane.sig.env.scoresdf,
  #   aes(
  #     x = 0,
  #     xend = NMDS1,
  #     y = 0,
  #     yend = NMDS2),
  #   arrow = arrow(length = unit(0.025, "npc")),
  #   size = 0.75
  # ) +
  #  geom_text(
#   data = subset(LowerMontane.sig.env.scoresdf),
#   check_overlap = T,
#   fontface = "italic",
#   size = 4,
#   nudge_x = 0.25,
#   nudge_y = -0.025,
#   vjust = 0.0,
#   #Pushes label away from the arrow
#   mapping = aes(
#     label = EnvPred2,
#     x = NMDS1 * 1.1,
#     y = NMDS2 * 1.1
#   )
# ) +
geom_abline(
  #Horizontal line through origin
  intercept = 0,
  slope = 0,
  linetype = "dashed",
  size = 0.5,
  colour = "gray"
) +
  geom_vline(
    #Vertical line through origin
    aes(xintercept = 0),
    linetype = "dashed",
    size = 0.5,
    colour = "gray"
  ) +
  
  # labs(title = "") +
  geom_text_repel(
    data = subset(fort_Families, Score == 'species'),
    # check_overlap = F,
    fontface = "bold",
    size = 3.5,
    color = "#51484f",
    # nudge_y = -0.025,
    #Pushes label away from the arrow
    mapping = aes(label = Label,
                  x = NMDS1,
                  y = NMDS2,)
  ) +
  theme_few() +
  theme(
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11),
    legend.position = "bottom",
    legend.key.width = unit(1, "cm")
  ) +
  labs(x = "NMS1", y = "NMS2")
# xlim (-1.5,1.5)+
# ylim(-1.5,1.5)

FamiliesBurnNMDSplot
```



```{r}
#Families NMDS plot
FamiliesForestNMDSplot <- ggplot() +
  geom_point(
    data = subset(fort.plot_Families, Score == "sites"),
    size = 2,
    alpha = 0.75,
    shape = "circle",
    stroke = 1,
    mapping = aes(x = NMDS1, y = NMDS2, color = forest_type)
  ) +
  scale_color_manual(
    name = "",
    labels = c("Lower Montane", "Upper Montane", "Subalpine"),
    values = c("#9ebc00", "#ffc107", "#981679")
  ) +
  # geom_segment(
  #   data = LowerMontane.sig.env.scoresdf,
  #   aes(
  #     x = 0,
  #     xend = NMDS1,
  #     y = 0,
  #     yend = NMDS2),
  #   arrow = arrow(length = unit(0.025, "npc")),
  #   size = 0.75
  # ) +
  #  geom_text(
#   data = subset(LowerMontane.sig.env.scoresdf),
#   check_overlap = T,
#   fontface = "italic",
#   size = 4,
#   nudge_x = 0.25,
#   nudge_y = -0.025,
#   vjust = 0.0,
#   #Pushes label away from the arrow
#   mapping = aes(
#     label = EnvPred2,
#     x = NMDS1 * 1.1,
#     y = NMDS2 * 1.1
#   )
# ) +
geom_abline(
  #Horizontal line through origin
  intercept = 0,
  slope = 0,
  linetype = "dashed",
  size = 0.5,
  colour = "gray"
) +
  geom_vline(
    #Vertical line through origin
    aes(xintercept = 0),
    linetype = "dashed",
    size = 0.5,
    colour = "gray"
  ) +
  geom_text_repel(
    data = subset(fort_Families, Score == 'species'),
    # check_overlap = F,
    fontface = "bold",
    size = 3.5,
    color = "#51484f",
    # nudge_y = 0,
    # vjust = 1,
    #Pushes label away from the arrow
    mapping = aes(label = Label,
                  x = NMDS1,
                  y = NMDS2,)
  ) +
  theme_few() +
  theme(
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11),
    legend.position = "bottom",
  ) +
  labs(x = "NMS1", y = "NMS2")
# xlim (-1.5,1.5)+
# ylim(-1.5,1.5)

FamiliesForestNMDSplot

```


```{r}

#Place plots side by side 

legend1 <- get_legend(FamiliesBurnNMDSplot)
legend2 <- get_legend(FamiliesForestNMDSplot)

legends = list(legend1, legend2)


ggarrange(FamiliesBurnNMDSplot, FamiliesForestNMDSplot)
  
# annotate_figure(Families_ggarrange, fig.lab = "NMDS1", fig.lab.pos = "bottom")
```

