---
title: "Thesis2022_Q1"
author: "Amanda Hastings"
date: "2022-09-19"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Q1: How are different tree species regenerating across forest types at low and high burn severities?

```{r}
#Read in data 
ThesisDataRaw <- read.csv("HastingsA_MasterThesisData8.8.22.csv")
TerraData <- read.csv("HastingsA_TerraclimateMaster8.16.csv")
ThesisSeedlingsData <- read.csv("HastingsA_MasterSeedlingsData9.12.22.csv")

#Load libraries
library(tidyverse)
library(dplyr)
library(RColorBrewer)
library(emmeans)
library(multcomp)
library(multcompView)
library(ggthemes)
library(gh)
library(pscl)
library(car)
```

## Formatting raw thesis site data and seedlings data 

```{r}
#Filter to only wanted sites in ThesisData 
ThesisData <- ThesisDataRaw %>% 
  filter(!PlotID %in% c('LS11','HS02','HS12','HS13', 'LS01_MV','LS02_MV','HS01_MV'))

ThesisSeedlings <- ThesisSeedlingsData %>% 
  filter(!PlotID %in% c('LS11','HS02','HS12','HS13'))%>% 
  dplyr::select(!c(Fire_name, regen_radius, regenplot_size_m2, regenplot_size_ha))%>% 
  select_if(~!any(is.na(.)))

```

## Combining site data with seedlings data

```{r}
ThesisAll1 <- full_join(ThesisData, ThesisSeedlings, by= "PlotID") %>% 
  arrange(PlotID)%>% 
  mutate(forest_type = factor(forest_type))

```

## Create data frame with na values for tree species not expected in forest types 

```{r}
ThesisSeedRename <- ThesisAll1 %>% 
  dplyr::select(forest_type, burn_sev_category, POTR_total, QUGA_total, RONE_total, PIPO_total, PSME_total, ABCO_total, PICO_total, PIFL_total, PIAR_total, ABLA_total, PIEN_total)%>%
  rename("POTR"= 3, "QUGA"= 4, "RONE"= 5, "PIPO"= 6, "PSME"= 7, "ABCO"= 8,  "PICO" = 9, "PIFL" = 10, "PIAR"=11, "ABLA"=12, "PIEN"=13)


ThesisNAs <- ThesisSeedRename %>%
  mutate(QUGA = replace(QUGA, forest_type == "Subalpine", NA))%>% 
  mutate(RONE = replace(RONE, forest_type == "Subalpine" | forest_type == "Upper Montane", NA))%>% 
  mutate(PIPO = replace(PIPO, forest_type == "Subalpine" | forest_type == "Upper Montane", NA))%>% 
  mutate(PIFL = replace(PIFL, forest_type == "Lower Montane", NA))%>% 
  mutate(PIAR = replace(PIAR, forest_type == "Lower Montane" | forest_type == "Upper Montane", NA))%>% 
  mutate(ABLA = replace(ABLA, forest_type == "Lower Montane" | forest_type == "Upper Montane", NA))%>% 
  mutate(PIEN = replace(PIEN, forest_type == "Lower Montane" | forest_type == "Upper Montane", NA))
  
         
```


# Seedlings long data
```{r}
# Covert seedlings data to long format
SeedlingsLong <- ThesisNAs %>%
  tidyr::pivot_longer(cols = c(POTR:PIEN),
                      names_to = "Species",
                      values_to = "Density") %>% 
  mutate(forest_type = factor(
    forest_type,
    levels = c("Lower Montane", "Upper Montane", "Subalpine")
  )) %>% 
  mutate(Species = factor(Species,
                          levels = c("PIPO","PSME","ABCO","PICO","PIFL", "PIAR","ABLA","PIEN", "QUGA", "RONE","POTR")))
```


# Facet axis removing labels for species not present in forest types 

```{r}
#Omit na values
SeedlingsLong2 <- SeedlingsLong %>% 
  mutate(Density = replace(Density, forest_type == "Subalpine" & Species == "PIEN", NA))%>% 
  mutate(Density = replace(Density, forest_type == "Subalpine" & Species == "ABLA", NA)) %>% 
  mutate(Density = replace(Density, forest_type == "Subalpine" & Species == "ABCO", NA)) %>%
  mutate(Density = replace(Density, forest_type == "Upper Montane" & Species == "ABCO", NA))%>% 
  mutate(Density = replace(Density, forest_type == "Lower Montane" & Species == "PIPO", NA)) %>%
  na.omit
```


# Seedlings plot 
```{r}
# Facet wrap by forest type - seedlings plot

ggplot(SeedlingsLong2,
       aes(x = Species, y = Density, fill = burn_sev_category)) +
  geom_boxplot() + 
  # ylim(0, 25000) +
  facet_wrap(~ forest_type, scales="free_x") +
  labs(x = "Species",
       y = "Seedlings per hectare",
       fill = "Burn Severity") +
  theme(legend.position = "bottom",
        legend.key.width = unit(1, "cm")) +
  theme_few() +
  scale_fill_manual(
    name = "Burn Severity",
    labels = c("High", "Low"), values=c("#bd0026","#fd8d3c"))+
  theme(
    axis.text.x = element_text(
      color = "black",
      size = 9,
      angle = 45,
      vjust = .8,
      hjust = .8
    ), 
    axis.title.x=element_text(size=12), 
    axis.title.y=element_text(size=12), 
    axis.text.y = element_text(size=10), 
    legend.title=element_text(size=12), 
    strip.text.x = element_text(size=12), 
    legend.text = element_text(size=11))
```

# Data frame analyses grouped by conifer/decid 

```{r}
#Add conifer and deciduous total columns 
#POTR_total, QUGA_total, RONE_total, PIPO_total, PSME_total, ABCO_total, PICO_total, PIFL_total, PIAR_total, ABLA_total, PIEN_total

decid_list = c("POTR_total", "QUGA_total", "RONE_total")
conifer_list = c("PIPO_total", "PSME_total", "ABCO_total", "PICO_total", "PIFL_total", "PIAR_total", "ABLA_total", "PIEN_total")


ThesisAll1 <- ThesisAll1 %>% 
  mutate("deciduous_total" = rowSums(ThesisAll1[,decid_list])) %>% 
  mutate("conifer_total" = rowSums(ThesisAll1[,conifer_list]))
```

# Conifer and deciduous data frame for plotting 

```{r}

DecidConiferLong <- ThesisAll1 %>%
  rename("Coniferous"= conifer_total)%>%
  rename("Deciduous"= deciduous_total)%>%
  dplyr::select(forest_type,
                burn_sev_category,
                Coniferous,
                Deciduous) %>%
  pivot_longer(
    cols = c("Coniferous", "Deciduous"),
    names_to = "TreeGroup",
    values_to = "Density"
  ) %>%
  mutate(forest_type = factor(
    forest_type,
    levels = c("Lower Montane", "Upper Montane", "Subalpine")
  ))
```



# Conifer and deciduous plot 

```{r}
ggplot(DecidConiferLong,
       aes(x = TreeGroup, y = Density, fill = burn_sev_category)) +
  geom_boxplot() + 
  # ylim(0, 25000) +
  facet_wrap(~ forest_type, scales="free_x") +
  labs(x = "Etymological Group",
       y = "Seedlings per hectare",
       fill = "Burn Severity") +
  theme(legend.position = "bottom",
        legend.key.width = unit(1, "cm")) +
  theme_few() +
  scale_fill_manual(
    name = "Burn Severity",
    labels = c("High", "Low"), values=c("#bd0026","#fd8d3c"))+
  theme(
    axis.text.x = element_text(
      color = "black",
      size = 10,
      angle = 0,
      vjust = 1.5,
      hjust = .5
    ), 
    axis.title.x=element_text(size=12), 
    axis.title.y=element_text(size=12), 
    axis.text.y = element_text(size=10), 
    legend.title=element_text(size=12), 
    strip.text.x = element_text(size=12), 
    legend.text = element_text(size=11))
```



# Summary statistics of conifer/decid seedlings densities across sites


```{r}
ConiferDecidSumStats <- ThesisAll1 %>% 
  group_by(forest_type, burn_sev_category)%>%
  summarize(conifer_mean = mean(conifer_total), 
            conifer_sd = sd(conifer_total), 
            deciduous_mean = mean(deciduous_total), 
            deciduous_sd = sd(deciduous_total)) 

length(which(ThesisAll1$conifer_total != 0.0000 & ThesisAll1$burn_sev_category == "Low"))
length(which(ThesisAll1$conifer_total != 0.0000 & ThesisAll1$burn_sev_category == "High"))
length(which(ThesisAll1$deciduous_total != 0.0000 & ThesisAll1$forest_type == "Lower Montane"))
length(which(ThesisAll1$deciduous_total != 0.0000 & ThesisAll1$forest_type == "Upper Montane"))
length(which(ThesisAll1$deciduous_total != 0.0000 & ThesisAll1$forest_type == "Subalpine"))

length(which(ThesisAll1$QUGA_total != 0.0000 & ThesisAll1$forest_type == "Lower Montane"))
length(which(ThesisAll1$QUGA_total != 0.0000 & ThesisAll1$forest_type == "Upper Montane"))
```


#------------Conifer/decid analyses---------------------------------------------


# Two way ANOVA deciduous density only, sqrt transform
```{r}
# Two way ANOVA deciduous, sqrt transform
ThesisAll1$deciduous_total<-as.integer(ThesisAll1$deciduous_total)

LMFitDecidsqrt <- lm(sqrt(deciduous_total) ~ burn_sev_category * forest_type, data = ThesisAll1)
anova(LMFitDecidsqrt)
par(mfrow = c(1, 2))
plot(LMFitDecidsqrt, which = 1:2)
```

# Tukey pairs Deciduous Density - Forest Type
```{r}
#Tukey pairs Deciduous density 
emoutDecid <- emmeans(LMFitDecidsqrt, ~forest_type)
pairs(emoutDecid)
cld(emoutDecid)
```


# Back transform deciduous seedlings

```{r}
backDecid <- emmeans(LMFitDecidsqrt, pairwise ~ forest_type, regrid = "response")
backDecid

```


## Conifer density 

# Conifer Poisson Regression
```{r}
ThesisAll1$conifer_total<-as.integer(ThesisAll1$conifer_total)

ConiferPoisson <- glm(conifer_total~ burn_sev_category*forest_type, data=ThesisAll1, family=poisson) 
summary(ConiferPoisson)
Anova(ConiferPoisson)
```


```{r}
emoutConifer <- emmeans(ConiferPoisson, ~burn_sev_category*forest_type, type= "response")
pairs(emoutConifer)
cld(emoutConifer)
```


#------------Seedlings and pre-fire basal area----------------------------------


# Conifer seedling presense/absense and pre-fire basal area 

```{r}
# PICO prefire basal area, seed source, seedlings
PICOSeedPrefireBA <- ThesisAll1 %>%
  mutate(
    "prefire_seed_PICO" = case_when(
      PICO_total != 0.0000 &
        (PICO_dead_ba != 0.0000 | PICO_live_ba != 0.0000) ~ "Success",
      PICO_total == 0.0000 &
        (PICO_dead_ba != 0.0000 | PICO_live_ba != 0.0000) ~ "Failure",
      PICO_total != 0.0000 &
        (PICO_dead_ba == 0.0000 & PICO_live_ba == 0.0000) ~ "New", 
      PICO_total == 0.0000 & PICO_dead_ba == 0.0000 & PICO_live_ba == 0.0000 ~ "NA")) %>% 
  dplyr::select(forest_type, burn_sev_category, PICO_total, PICO_dead_ba, PICO_live_ba, PICO_seed_source, prefire_seed_PICO)
 
 
# ABCO prefire basal area, seed source, seedlings
ABCOSeedPrefireBA <- ThesisAll1 %>%
  mutate(
    "prefire_seed_ABCO" = case_when(
      ABCO_total != 0.0000 &
        (ABCO_dead_ba != 0.0000 | ABCO_live_ba != 0.0000) ~ "Success",
      ABCO_total == 0.0000 &
        (PICO_dead_ba != 0.0000 | ABCO_live_ba != 0.0000) ~ "Failure",
      ABCO_total != 0.0000 &
        (ABCO_dead_ba == 0.0000 & ABCO_live_ba == 0.0000) ~ "New", 
      ABCO_total == 0.0000 & ABCO_dead_ba == 0.0000 & ABCO_live_ba == 0.0000 ~ "NA")) %>% 
  dplyr::select(forest_type, burn_sev_category, ABCO_total, ABCO_dead_ba, ABCO_live_ba, ABCO_seed_source, prefire_seed_ABCO)
  

# PSME prefire basal area, seed source, seedlings
PSMESeedPrefireBA <- ThesisAll1 %>%  
   mutate(
    "prefire_seed_PSME" = case_when(
      PSME_total != 0.0000 &
        (PSME_dead_ba != 0.0000 | PSME_live_ba != 0.0000) ~ "Success",
      PSME_total == 0.0000 &
        (PSME_dead_ba != 0.0000 | PSME_live_ba != 0.0000) ~ "Failure",
      PSME_total != 0.0000 &
        (PSME_dead_ba == 0.0000 & PSME_live_ba == 0.0000) ~ "New", 
      PSME_total == 0.0000 & PSME_dead_ba == 0.0000 & PSME_live_ba == 0.0000 ~ "NA"))%>% 
  dplyr::select(forest_type, burn_sev_category, PSME_total, PSME_dead_ba, PSME_live_ba, PSME_seed_source, prefire_seed_PSME)
  
   
# PIFL prefire basal area, seed source, seedlings
PIFLSeedPrefireBA <- ThesisAll1 %>%
   mutate(
    "prefire_seed_PIFL" = case_when(
      PIFL_total != 0.0000 &
        (PIFL_dead_ba != 0.0000 | PIFL_live_ba != 0.0000) ~ "Success",
      PIFL_total == 0.0000 &
        (PIFL_dead_ba != 0.0000 | PIFL_live_ba != 0.0000) ~ "Failure",
      PIFL_total != 0.0000 &
        (PIFL_dead_ba == 0.0000 & PIFL_live_ba == 0.0000) ~ "New", 
      PIFL_total == 0.0000 & PIFL_dead_ba == 0.0000 & PIFL_live_ba == 0.0000 ~ "NA")) %>%
  dplyr::select(forest_type, burn_sev_category, PIFL_total, PIFL_dead_ba, PIFL_live_ba, PIFL_seed_source, prefire_seed_PIFL)


# PIAR prefire basal area, seed source, seedlings
PIARSeedPrefireBA <- ThesisAll1 %>%
     mutate(
    "prefire_seed_PIAR" = case_when(
      PIAR_total != 0.0000 &
        (PIAR_dead_ba != 0.0000 | PIAR_live_ba != 0.0000) ~ "Success",
      PIAR_total == 0.0000 &
        (PIAR_dead_ba != 0.0000 | PIAR_live_ba != 0.0000) ~ "Failure",
      PIAR_total != 0.0000 &
        (PIAR_dead_ba == 0.0000 & PIAR_live_ba == 0.0000) ~ "New", 
      PIAR_total == 0.0000 & PIAR_dead_ba == 0.0000 & PIAR_live_ba == 0.0000 ~ "NA")) %>%
  dplyr::select(forest_type, burn_sev_category, PIAR_total, PIAR_dead_ba, PIAR_live_ba, PIAR_seed_source, prefire_seed_PIAR)

```


# Deciduous seedling presense/absense and pre-fire basal area 

```{r}
# POTR prefire basal area, seed source, seedlings
POTRSeedPrefireBA <- ThesisAll1 %>%  
   mutate(
    "prefire_seed_POTR" = case_when(
      POTR_total != 0.0000 &
        (POTR_dead_ba != 0.0000 | POTR_live_ba != 0.0000) ~ "Success",
      POTR_total == 0.0000 &
        (POTR_dead_ba != 0.0000 | POTR_live_ba != 0.0000) ~ "Failure",
      POTR_total != 0.0000 &
        (POTR_dead_ba == 0.0000 & POTR_live_ba == 0.0000) ~ "New", 
      POTR_total == 0.0000 & POTR_dead_ba == 0.0000 & POTR_live_ba == 0.0000 ~ "NA"))%>% 
  dplyr::select(forest_type, burn_sev_category, POTR_total, POTR_dead_ba, POTR_live_ba, POTR_seed_source, prefire_seed_POTR)
  
   
# # QUGA prefire basal area, seed source, seedlings
# QUGASeedPrefireBA <- ThesisAll1 %>%
#    mutate(
#     "prefire_seed_QUGA" = case_when(
#       QUGA_total != 0.0000 &
#         (QUGA_dead_ba != 0.0000 | QUGA_live_ba != 0.0000) ~ "Success",
#       QUGA_total == 0.0000 &
#         (QUGA_dead_ba != 0.0000 | QUGA_live_ba != 0.0000) ~ "Failure",
#       QUGA_total != 0.0000 &
#         (QUGA_dead_ba == 0.0000 & QUGA_live_ba == 0.0000) ~ "New", 
#       QUGA_total == 0.0000 & QUGA_dead_ba == 0.0000 & QUGAL_live_ba == 0.0000 ~ "NA")) %>%
#   dplyr::select(forest_type, burn_sev_category, QUGA_total, QUGA_dead_ba, QUGA_live_ba, QUGA_seed_source, prefire_seed_QUGA)


# # RONE prefire basal area, seed source, seedlings
# RONESeedPrefireBA <- ThesisAll1 %>%
#      mutate(
#     "prefire_seed_RONE" = case_when(
#       RONE_total != 0.0000 &
#         (PIAR_dead_ba != 0.0000 | RONE_live_ba != 0.0000) ~ "Success",
#       RONE_total == 0.0000 &
#         (PIAR_dead_ba != 0.0000 | RONE_live_ba != 0.0000) ~ "Failure",
#       RONE_total != 0.0000 &
#         (PIAR_dead_ba == 0.0000 & RONE_live_ba == 0.0000) ~ "New", 
#       RONE_total == 0.0000 & RONE_dead_ba == 0.0000 & RONE_live_ba == 0.0000 ~ "NA")) %>%
#   dplyr::select(forest_type, burn_sev_category, RONE_total, RONE_dead_ba, RONE_live_ba, RONE_seed_source, prefire_seed_RONE)
```



#------------Deciduous speices only analyses------------------------------------


# Two way ANOVA QUGA density only, sqrt transform
```{r}
# Two way ANOVA deciduous, sqrt transform
ThesisAll1$QUGA_total<-as.integer(ThesisAll1$QUGA_total)


QUGAdensity <- ThesisAll1 %>% 
  dplyr::select(PlotID, aspect_rescaled, heat_load, field.elevation..m., burn_sev_category, forest_type, QUGA_total)%>% 
  filter(!forest_type == "Subalpine")


hist(sqrt(QUGAdensity$QUGA_total))

QUGALMFit <- lm(sqrt(QUGA_total) ~ burn_sev_category+forest_type, data = QUGAdensity)
anova(QUGALMFit)
par(mfrow = c(1, 2))
plot(QUGALMFit, which = 1:2)


QUGAPoisson <- glm(QUGA_total~ burn_sev_category+forest_type, data=QUGAdensity, family=poisson) 
summary(QUGAPoisson )
Anova(QUGAPoisson)
```


```{r}
ggplot(QUGAdensity,
       aes(x = heat_load, y = QUGA_total)) +
  geom_point()

cor.test(QUGAdensity$QUGA_total, QUGAdensity$heat_load)     
```


```{r}
emoutQUGAPoisson <- emmeans(QUGAPoisson, ~burn_sev_category*forest_type, type= "response")
pairs(emoutQUGAPoisson)
cld(emoutQUGAPoisson)
```


# Two way ANOVA POTR density only, sqrt transform
```{r}
# Two way ANOVA deciduous, sqrt transform
ThesisAll1$POTR_total<-as.integer(ThesisAll1$POTR_total)

hist(sqrt(ThesisAll1$POTR_total))


POTRLMFit <- lm(sqrt(POTR_total) ~ burn_sev_category*forest_type, data = ThesisAll1)
anova(POTRLMFit)
par(mfrow = c(1, 2))
plot(POTRLMFit, which = 1:2)
```

# Test of correlation for POTR densities and heat load 

```{r}
ggplot(ThesisAll1,
       aes(x = heat_load, y = POTR_total)) +
  geom_point()


cor.test(ThesisAll1$POTR_total, ThesisAll1$heat_load)
```


# One-way ANOVA RONE density only, 

```{r}
# Two way ANOVA deciduous, sqrt transform
ThesisAll1$RONE_total<-as.integer(ThesisAll1$RONE_total)


RONEdensity <- ThesisAll1 %>% 
  dplyr::select(PlotID, aspect_rescaled, heat_load, field.elevation..m., burn_sev_category, forest_type, RONE_total)%>% 
  filter(forest_type == "Lower Montane")


hist(sqrt(RONEdensity$RONE_total))

RONELMFit <- lm(sqrt(RONE_total) ~ burn_sev_category, data = RONEdensity)
anova(RONELMFit)
par(mfrow = c(1, 2))
plot(RONELMFit, which = 1:2)

cor.test(RONEdensity$RONE_total, RONEdensity$heat_load)
```



# Correlation conifers 

```{r}
LowerMontane <- ThesisAll1 %>% 
  filter(forest_type == "Lower Montane")

cor.test(LowerMontane$conifer_total, LowerMontane$heat_load)
```

# PICO only 

```{r}
ThesisAll1$PICO_total<-as.integer(ThesisAll1$PICO_total)

hist(sqrt(ThesisAll1$PICO_total))


PICOLMFit <- lm(sqrt(PICO_total) ~ burn_sev_category*forest_type, data = ThesisAll1)
anova(PICOLMFit)
par(mfrow = c(1, 2))
plot(PICOLMFit, which = 1:2)
```









